The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-04-02 src/ms.tex
supporting maintainers

Sphinx is a common and useful tool for generating documentation for Python
packages. The Astropy project maintains a default Sphinx configuration along
with Astropy-specific extensions that can be easily added to community
projects via the
\href{https://github.com/astropy/sphinx-astropy}{\texttt{sphinx-astropy}} meta
package. This tool provides a pre-configured Sphinx setup compatible with
Astropy projects, which includes several extensions useful for generating API
documentation
(\href{https://github.com/astropy/sphinx-automodapi}{\texttt{sphinx-automodapi}}),
allowing for Numpy docstring parsing
(\href{https://github.com/numpy/numpydoc}{\texttt{numpydoc}}), embedded image
handling
(\href{https://github.com/sphinx-gallery/sphinx-gallery}{\texttt{sphinx-gallery}};
\href{https://github.com/python-pillow}{\texttt{pillow}}), advanced
documentation testing support
(\href{https://github.com/astropy/pytest-doctestplus}{\texttt{pytest-doctestplus}}),
and providing a custom documentation theme ideal for analysis packages
(\href{https://github.com/astropy/sphinx-astropy}{\texttt{astropy-sphinx-theme}}).

@@

Over the last decade, the needs of (and challenges faced by) the \astropy
Project have evolved, but many common themes persist.
For example, ...list...
In many cases, these themes are not unique to \astropy: Many Open-Source
Software communities are facing and addressing similar issues.
Below we discuss a few of these challenges and ...

\begin{description}
  \item[Supporting an active group of maintainers] With funding, can pay maintainers. But: time commitment, some positions don't allow. Need for new career paths on the spectrum from software engineer to astronomy researcher. Also need to attract new non-hired contributors to sustain the project, but challenging when the code has become quite complex (astropy as an "institution", don't want to be a walled-off group). (Need for fast feedback and code reviews to empower contributors, but puts a burden on maintainers)
  \item[Long-term / sustained funding for maintaining infrastructure] Further from "science," still critical to astropy, but who will provide updates and maintenance? More than just updating versions and releasing. (grants last 3 years; infrastructure evolves, nature of open source ecosystem)
  \item[Understanding, communicating with, and engaging our user communities] Different overlapping and intersecting communities of users: telescopes, archives, researchers, students, educators, general public. Communities have different needs from Astropy. Also, as ecosystem grows, making sure users know about all of the functionality (workshops, tutorials, educational materials, discoverability, accessibility) (Continued engagement with user community).
\end{description}

>>>>

Sphinx is a common and useful tool for generating documentation for Python
packages. The Astropy project maintains a default Sphinx configuration along
with Astropy-specific extensions that can be easily added to community
projects via the
\href{https://github.com/astropy/sphinx-astropy}{\texttt{sphinx-astropy}} meta
package. This tool provides a pre-configured Sphinx setup compatible with
Astropy projects, which includes several extensions useful for generating API
documentation
(\href{https://github.com/astropy/sphinx-automodapi}{\texttt{sphinx-automodapi}}),
allowing for \package{numpy} docstring parsing
(\href{https://github.com/numpy/numpydoc}{\texttt{numpydoc}}), embedded image
handling
(\href{https://github.com/sphinx-gallery/sphinx-gallery}{\texttt{sphinx-gallery}};
\href{https://github.com/python-pillow}{\texttt{pillow}}), advanced
documentation testing support
(\href{https://github.com/astropy/pytest-doctestplus}{\texttt{pytest-doctestplus}}),
and providing a custom documentation theme ideal for analysis packages
(\href{https://github.com/astropy/sphinx-astropy}{\texttt{astropy-sphinx-theme}}).

@@

Over the history of \astropy, the needs and challenges of the \astropy Project
have evolved, but many common themes have persisted and are shared by other
open source software communities.
However, \astropy also faces a number of unique challenges stemming from the
fact that the Project is in a special position in the astronomy software stack:
The \astropy code and ecosystem are more specialized than core scientific tools
like \package{numpy} and \package{scipy}, but more general than software
packages developed to enable specific research subdomains like
\package{exoplanet} \citep{exoplanet:2021}.
Below we discuss a few challenges faced by the Project in more detail, but we
stress that this list is not complete or exhaustive.
\begin{description}
  \item[Supporting an active group of maintainers] Throughout its history,
  development and maintenance of code and educational materials in the \astropy
  Project --- the \astropypkg core package, coordinated packages, Learn Astropy
  repositories, and Project infrastructure packages --- has been done by a mix
  of volunteer effort and institutional support.
  At present, the \href{https://www.astropy.org/team.html}{named roles} and
  maintainer positions in the \astropy Project are also split roughly evenly
  between these types of contributors.
  As is the case with many other open-source software communities, the culture
  of the Project and overall coordination relies heavily on volunteer effort,
  but we therefore have significant amounts of critical infrastructure or core
  functionality that is either maintained with no redundancy or unmaintained as
  contributors naturally move on from and stop contributing to the Project.
  Within the core package, a prime example is the \astropyfits subpackage, which
  is largely in maintenance mode and maintained by just one contributor (despite
  the significant codebase and user community behind this subpackage).
  While we now have funding to support the longer-term sustainability of the
  Project and thus to address these issues, we are finding it difficult to use
  this funding in ways that address our needs.
  For one, many of the needs of the Project require long-term and sustained
  funding rather than short-term contracts.
  But also the specialized nature of the software requires being able to hire or
  contract software engineers with domain knowledge about astronomical software
  tools.
  Finally, we want to be able to preserve and sustain the engaged community of
  volunteers, software-engaged researchers, and existing software engineers that
  contribute to the Project and that have enabled its success in ways that
  support their careers.
  These challenges will require cultural and larger systemic changes to
  astronomical research and career paths, and so will take time to influence and
  implement.
  \item[Diversity of users, contributors, maintainers, and coordinators]
  Astropy is
  Practical issues: translation of content and documentation, github is not accessible in some countries.
  \item[Long-term / sustained funding for maintaining infrastructure] Software
  and packaging infrastructure includes but not limited to Continuous
  Integration (CI), test plugins, documentation builds, and package delivery.
  Keeping the infrastructure working is essential in software development and
  operations. This kind of work has low visibility and is not attractive to
  scientists or volunteer contributors; therefore, it is most effective to
  delegate such tasks to paid roles. \astropy secured a three-year grant for
  this but the situation remains uncertain once the grant runs out, as
  infrastructure evolves and requires constant maintenance and upgrades. Some of
  the engineers currently in these roles are also employees of research
  institutions that may leave the Project without much notice, adding to the
  uncertainty factor.
  \item[Understanding, communicating with, and engaging our user communities] Different overlapping and intersecting communities of users: telescopes, archives, researchers, students, educators, general public. Communities have different needs from Astropy. Also, as ecosystem grows, making sure users know about all of the functionality (workshops, tutorials, educational materials, discoverability, accessibility) (Continued engagement with user community).
\end{description}
